Computationally Efficient Simulation of Queues: The \proglang{R} Package \pkg{queuecomputer}

### 2018-11-06 main.tex
Review use of commas.

%% an abstract and keywords
\Abstract{
Large networks of queueing systems model important real-world systems such as MapReduce clusters, web-servers, hospitals, call-centers and airport passenger terminals. To model such systems accurately we must infer queueing parameters from data. Unfortunately, for many queueing networks there is no clear way to proceed with parameter inference from data. Approximate Bayesian computation could offer a straight-forward way to infer parameters for such networks if we could simulate data quickly enough.

@@

The queues we encounter in our everyday experience, where customers wait in line to be served by a server, are a useful analogy for many other processes. We say analogy because the word customers could represent: MapReduce jobs \citep{lin_joint_2013}; patients in a hospital \citep{takagi_queueing_2016}; items in a manufacturing system \citep{dallery_manufacturing_1992}; calls to a call center \citep{gans_telephone_2003}; shipping containers in a seaport \citep{kozan_comparison_1997} or even cognitive tasks \citep{cao_queueing_2013}. Similarly server could represent: a compute cluster; medical staff; machinery or a customer service representative at a call center. Queueing systems can also be networked together to form queueing networks. We can use queueing networks to build models of processes such as provision of internet services \citep{sutton_bayesian_2011}, passenger facilitation at international airports \citep{wu_review_2013} and emergency evacuations \citep{van_woensel_modeling_2007}. Clearly queueing systems and queueing networks are useful for understanding important real-world systems.

Performance measures for a given queueing system can often only be derived through simulation. Queues are usually simulated with discrete event simulation (DES) \citep[pg. 226]{insua2012bayesian}. In DES changes in state are discontinuous. The state is acted upon by a countable list of events at certain times which cause the discontinuities. If the occurrence of an event is independent of everything except simulation time it is determined, otherwise it is contingent \citep{nance1981time}.

@@

\pkg{queuecomputer} \citep{Rpkg_queuecomputer} implements an algorithm that can easily be applied to a wide range of queueing systems and networks of queueing systems. It is vastly more computationally efficient than existing approaches to DES. We term this new computationally efficient algorithm queue departure computation (QDC). Computational efficiency is important because if we can simulate from queues quickly then we can embed a queue simulation within an approximate Bayesian computation (ABC) algorithm \citep{sunnaker_approximate_2013} and estimate queue parameters for very complicated queueing models in a straight forward manner.

@@

The simplest queueing system is exponential in distribution for both the inter-arrival $\delta_i \overset{iid}{\sim} \mathrm{exp}(\lambda) ~ \forall i \in 1:n$ and service processes $s_i \overset{iid}{\sim} \mathrm{exp}(\mu) ~ \forall i \in 1:n$, where $\lambda$ and $\mu$ are exponential rate parameters. Additionally, $K$ is set to 1, $C$ and $n$ are infinite and $R$ is FCFS. It is denoted by $M/M/1/ \infty / \infty / FCFS$, which is shortened to $M/M/1$.

@@

In the case of a $M/M/K$ system equilibrium distributions for performance measures are derived analytically, they are found in standard queueing theory textbooks \citep{lipsky2008queueing, thomopoulos2012fundamentals}. For instance the limit probability of $N$ customers in the system $\Prob(N)$  is

@@

If the parameters of $f_\delta$ and $f_{\mathbf{s}}$ are uncertain then we must turn to predictive distributions for estimates of performance measures, where are computed analytically for M/M/K queues (Equations \ref{eq:en} and \ref{eq:ew}). Predictive distributions of performance measures using Bayesian posterior distributions are derived by \citet{armero1994bayesian, armero1999dealing}.

\citet{jackson_networks_1957} was one of the first to consider networks of queueing systems. In a Jackson network there is a set of $J$ queueing systems. After a customer is served by queueing system $j$ they arrive at another queueing system with fixed probability $p_{j,k}$. Customers leave the system with probability $1 - \sum_{k=1}^{J} p_{j,k}$. Other examples of queueing networks include the tandem \citep{glynn_departures_1991}, parallel \citep{hunt_fast_1995} and the fork/join \citep{kim_analysis_1989} topologies.

@@

Most models of queueing systems assume time-invariant inter-arrival and service processes. In practice many real-world queues have inter-arrival processes which are strongly time-dependent, such as: call-centers \citep{weinberg2007bayesian, brown2005statistical}, airport runways \citep{koopman1972air} and hospitals \citep{brahimi_queueing_1991}. In the case of the $M/M/1$ queue we can adapt the notation to $M(t)/M(t)/1$ to represent exponential  processes where parameters $\lambda(t)$ and $\mu(t)$ change with time. Such queueing systems are referred to as dynamic queueing systems.

@@

QDC can be considered as a multiserver extension to an algorithm presented by \citet{lindley_theory_1952}. For a single server queueing system, the departure time of the $i$th customer is: $d_i = \max{(a_i, d_{i-1})} + s_i$, since the customer either waits for a server or the server waits for a customer. The algorithm (not the paper) was, surprisingly, not extended to multiserver systems until \citet{krivulin_recursive_1994}. However with each new customer $i$ the algorithm must search a growing $i+1$ length vector. This algorithm therefore scales poorly, with computational complexity $O(n^2)$, where $n$ is number of customers. \citet{kin_generalized_2010} adapted the original algorithm of \citet{kiefer_theory_1955} to an $O(nK)$ algorithm for multiserver tandem queues with blocking, that is $G/G/K/C$ queueing systems where $C$ is the maximum capacity number of customers in the queueing systems.

QDC can also be viewed as a computationally efficient solution to the set of equations presented in \citet[pg. 259]{sutton_bayesian_2011} for FCFS queueing systems. There is a single queue served by a fixed number of $K$ servers. The $i$th customer observes a set of times $b_i = \{ b_{ik} | k \in 1:K \}$ which represents the times when each server will next be available. The customer $i$ selects the earliest available server $p_i = \text{argmin}(b_i)$ from $b_i$. The departure time for the $i$th customer is therefore $d_i = \max({a_i, b_{p_i}}) + s_i$, since the server must wait for the customer or the customer must wait for the server. The QDC algorithm for a fixed number of servers (Algorithm \ref{alg:fixed_servers}) pre-sorts the arrival times. Rather than assigning a $b_i$ for each customer $i$ to form the matrix $\mathbf{b} \in \mathbf{M}^{n \times K}$, QDC considers $\mathbf{b}$ as a continually updated $K$ length vector representing the state of the system.

@@

This algorithm is simple and computationally efficient. At each iteration of the loop we need only search $\mathbf{b}$, a $K$ length vector for the minimum element in code line \ref{alg_line:search}. In the language of DES, we consider $\mathbf{b}$ as the system state and $\mathbf{a}$ as the event list, which are all determined events. This differs from conventional DES approaches to modelling queueing systems where the queue length is the system state and both $\mathbf{a}$ and $\mathbf{d}$ constitute the event list, where the events of $\mathbf{a}$ are determined and the events of $\mathbf{d}$ are continually updated and therefore contingent.

@@

We close server $k$ by writing an $\infty$ symbol to $b_k$ ensuring that no customer can use that server. If the server needs to be open again at time $t$ we write $t$ to $b_k$ allowing customers to use that server. Since $\mathbf{x}$ now corresponds to changes in $\mathbf{b}$ it is part of the event list along with $\mathbf{a}$. The entire event list is still determined and need not be updated mid-simulation.

@@

This algorithm can simulate queues of form $G(t)/G(t)/K(t)/\infty/n/FCFS$, where $K(t)$ refers to the number of open servers changing with time. As mentioned previously this algorithm is subject to Condition \ref{eq:condition}, this condition is not overly restrictive if we consider realistic systems with few changes in $K$. The recorded server allocations $\mathbf{p} = (p_1, \cdots, p_n)$ may not reflect the real system since Algorithm \ref{alg:changing_servers} does not allow the user to specify exactly which servers are open in each epoch, only how many are open and closed. If this output is needed or in cases where Condition \ref{eq:condition} does not hold, we must use the less computationally efficient but more general unconditional algorithm below.

@@

If Condition \ref{eq:condition} does not hold or if otherwise we wish to control exactly which servers are open at what time then we must use a less computationally efficient algorithm (Algorithm \ref{alg:changing_servers_unconditional}). Each server $k$ has its own partition of $L_k$ knot locations $\mathbf{x_k} = (x_{k,1}, \cdots, x_{k,L_k} ) \in \mathbb{R}^{L_k}_{+}$ and each $\mathbf{y_k} = (y_{k,1}, \cdots, y_{k,L_k+1} )$ is an alternating sequence of 0 and 1s of length $L_k$ indicating whether the server is open or closed respectively for the associated epoch. The vector $\mathbf{c}$ is used slightly differently to how it is used in \citet{sutton_bayesian_2011}. We use it to represent the time at which each server is next available for the current customer $i$, given the current system state $\mathbf{b}$. It is the output of the \code{next\_fun} function.

@@

This algorithm can simulate queueing systems of form $G(t)/G(t)/K(t)/\infty/n/FCFS$, where $K(t)$ refers to the number of open servers changing with time. In addition we can specify which particular servers are available when, not just how many and we are not bound by Condition \ref{eq:condition}. Once again we note that $\mathbf{b}$ can be considered as the system state and the event list is formed by $\mathbf{a}$ and the elements of $\mathbf{\underline{x}}$. This function can be called with the \code{queue\_step} function in \pkg{queuecomputer} by supplying a \code{server.list} object to the \code{servers} argument. For the rest of this paper we focus on Algorithms \ref{alg:fixed_servers} and \ref{alg:changing_servers} for their relative conceptual simplicity and computational efficiency.

@@

With the algorithms so far presented we can simulate from a very general set of queueing systems $G(t)/G(t)/K(t)/\infty/n/FCFS$ in a computationally efficient manner. In contrast to the algorithm of \cite{kin_generalized_2010}, the state vector $\mathbf{b}$ is written over in each iteration. The memory usage for QDC therefore scales with $O(n)$ rather than $O(nK)$.

@@

If the last element of $\mathbf{y}$ is zero it is possible that some customers will never be served, this is the ``Missed customers'' output. The performance measures that follow are the mean waiting time $\bar{w}$, the mean response time $\bar{r} = d - a$, the observed utilization factor $\bar{B}/K$, the mean queue length and the mean number of customers in the system respectively. The utilization factor $\bar{B}/K$ takes into account the changing number of open servers $K(t)$ where Algorithm \ref{alg:changing_servers} is used. We now explain the implementation details of package.

@@

The \code{queue\_step} calls the more primitive \code{queue} function which is a wrapper for S3 methods which implement Algorithms \ref{alg:fixed_servers}, \ref{alg:changing_servers} or \ref{alg:changing_servers_unconditional} depending on the class of the object supplied to the \code{server} argument of \code{queue\_step}. If \code{class(server)} is numeric then \code{queue} runs Algorithm \ref{alg:fixed_servers}, if it is a \code{server.stepfun} then \code{queue} runs Algorithm \ref{alg:changing_servers}, if it is a \code{server.list} then \code{queue} runs Algorithm \ref{alg:changing_servers_unconditional}. The \code{queue} function computes departure times $\mathbf{d}$ and server allocations $\mathbf{p}$ and the \code{queue\_step} function adds additional output such as waiting times and queue lengths which are used in summary and plot methods.

To simulate fork/join networks the \pkg{queuecomputer} function \code{wait\_step} provides a simple wrapper to the \pkg{base} function \code{pmax.int}, this function computes the maximum of each row for a set of two equal length numeric vectors. The vectors represent the departure times for each subjob and the departure time for the entire job is the maximum of each subjob.

@@

To demonstrate the validity of the algorithm we consider a $M/M/2/\infty/1000/FCFS$ queue. If QDC is valid for any $M/M/K$ queueing system then it is valid for any $G(t)/G(t)/K$ queueing system. This is because any non-zero $\mathbf{(a,s)}$ could conceivably come from two exponential distributions, even if the probability of the particular realization is vanishingly small. We replicate exact departure times computed with the \pkg{simmer} and \pkg{simpy} packages using \pkg{queuecomputer}. First we generate $\mathbf{a}$ and $\mathbf{s}$ to be used as input to all three packages.

@@

We now input these objects into the three scripts using \pkg{queuecomputer}, \pkg{simmer}, or \pkg{simpy}. First we run the \pkg{queuecomputer} script. The \code{queuecomputer_output} object is sorted in ascending order so that the departure times can be compared to the DES packages.

@@

Clearly QDC and its implementation \pkg{queuecomputer} are a more computationally efficient way to simulate queueing systems of the form $G(t)/G(t)/K/\infty/M/FCFS$ than conventional DES algorithms implemented by \pkg{simpy} and \pkg{simmer}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width = 0.7\textwidth]{figures/benchmark_083.pdf}
  \caption{Computation time in milliseconds for varying numbers of passengers for each DES/queueing package. Each package returns exactly the same set of departure times, since the same arrival and service times are supplied. The computation time reported here is the median time of 100 runs for each number of customers and each package. Intel (R) Core(TM) i7-6700
CPU @ 3.40GHz running Debian GNU/Linux.
 }
  \label{fig:bm_numberofpassengers}
\end{figure}

@@

We demonstrate \pkg{queuecomputer} by simulating a call center. The arrival time for each customer is the time that they called, the service time is how long it takes for their problem to be resolved once they reach an available customer service representative. Let's assume that the customers arrive by a homogeneous Poisson process over the course of the day.

@@

Dashawn however had to wait for an available server, since he arrived after the first two customers arrived but before the first two customers departed. He must wait until one of these customers departs before he can be served. We add the departure time of the first customer (Beatriz) to his service time to compute his departure time.

@@

So the first two customers had no waiting time but Dashawn had to wait for an available server. We can compute the waiting times for all three customers in this manner:

@@

The package integrates naturally with the popular data manipulation \proglang{R} package \pkg{dplyr} \citep{Rpkg_dplyr}. We demonstrate how to integrate \pkg{queuecomputer} and \pkg{dplyr} with a more complex Airport Terminal than before (Figure~\ref{fig:larger_airport}). Passengers from a set of 120 flights disembark at the arrivals concourse, and proceed through immigration using either the ``smart gate" or the ``manual gate" route, we therefore have two queues in parallel. The route taken (smart gate or manual gate) by each passenger is predetermined, but the server used by the passenger within these separate queueing systems is not.

Their bags are unloaded from the flights and proceed to the baggage hall with a delay, the division of a passengers and bags is a fork/join network. The bags and passengers are forked at the arrival concourse and joined at the baggage hall. After immigration the passengers proceed on to the baggage hall where they pick up their bags.

\begin{figure}[!htb]
\centering
\includegraphics[width = 0.8\textwidth]{figures/Larger_airport.pdf}
\caption{Diagram of larger airport scenario, there are 120 flights in total and two multiserver queueing systems operate in parallel. Passengers are preassigned to travel through either the ``manual" or ``smart gate" route through immigration. The passengers and bags are ``forked'' when each aircraft arrives are are ``joined'' at the baggage hall. }
\label{fig:larger_airport}
\end{figure}

>>>>

%% an abstract and keywords
\Abstract{
Large networks of queueing systems model important real-world systems such as MapReduce clusters, web-servers, hospitals, call-centers and airport passenger terminals. To model such systems accurately, we must infer queueing parameters from data. Unfortunately, for many queueing networks there is no clear way to proceed with parameter inference from data. Approximate Bayesian computation could offer a straight-forward way to infer parameters for such networks if we could simulate data quickly enough.

@@

The queues we encounter in our everyday experience, where customers wait in line to be served by a server, are a useful analogy for many other processes. We say analogy because the word customers could represent: MapReduce jobs \citep{lin_joint_2013}; patients in a hospital \citep{takagi_queueing_2016}; items in a manufacturing system \citep{dallery_manufacturing_1992}; calls to a call center \citep{gans_telephone_2003}; shipping containers in a seaport \citep{kozan_comparison_1997} or even cognitive tasks \citep{cao_queueing_2013}. Similarly, server could represent: a compute cluster; medical staff; machinery or a customer service representative at a call center. Queueing systems can also be networked together to form queueing networks. We can use queueing networks to build models of processes such as provision of internet services \citep{sutton_bayesian_2011}, passenger facilitation at international airports \citep{wu_review_2013} and emergency evacuations \citep{van_woensel_modeling_2007}. Clearly queueing systems and queueing networks are useful for understanding important real-world systems.

Performance measures for a given queueing system can often only be derived through simulation. Queues are usually simulated with discrete event simulation (DES) \citep[pg. 226]{insua2012bayesian}. In DES changes in state are discontinuous. The state is acted upon by a countable list of events at certain times which cause the discontinuities. If the occurrence of an event is independent of everything except simulation time it is determined; otherwise, it is contingent \citep{nance1981time}.

@@

\pkg{queuecomputer} \citep{Rpkg_queuecomputer} implements an algorithm that can easily be applied to a wide range of queueing systems and networks of queueing systems. It is vastly more computationally efficient than existing approaches to DES. We term this new computationally efficient algorithm queue departure computation (QDC). Computational efficiency is important because if we can simulate from queues quickly, then we can embed a queue simulation within an approximate Bayesian computation (ABC) algorithm \citep{sunnaker_approximate_2013} and estimate queue parameters for very complicated queueing models in a straightforward manner.

@@

The simplest queueing system is exponential in distribution for both the inter-arrival $\delta_i \overset{iid}{\sim} \mathrm{exp}(\lambda) ~ \forall i \in 1:n$ and service processes $s_i \overset{iid}{\sim} \mathrm{exp}(\mu) ~ \forall i \in 1:n$, where $\lambda$ and $\mu$ are exponential rate parameters. Additionally, $K$ is set to 1, $C$ and $n$ are infinite, and $R$ is FCFS. It is denoted by $M/M/1/ \infty / \infty / FCFS$, which is shortened to $M/M/1$.

@@

In the case of a $M/M/K$ system equilibrium distributions for performance measures are derived analytically, they are found in standard queueing theory textbooks \citep{lipsky2008queueing, thomopoulos2012fundamentals}. For instance, the limit probability of $N$ customers in the system $\Prob(N)$  is

@@

If the parameters of $f_\delta$ and $f_{\mathbf{s}}$ are uncertain, then we must turn to predictive distributions for estimates of performance measures, where are computed analytically for M/M/K queues (Equations \ref{eq:en} and \ref{eq:ew}). Predictive distributions of performance measures using Bayesian posterior distributions are derived by \citet{armero1994bayesian, armero1999dealing}.

\citet{jackson_networks_1957} was one of the first to consider networks of queueing systems. In a Jackson network, there is a set of $J$ queueing systems. After a customer is served by queueing system $j$, they arrive at another queueing system with fixed probability $p_{j,k}$. Customers leave the system with probability $1 - \sum_{k=1}^{J} p_{j,k}$. Other examples of queueing networks include the tandem \citep{glynn_departures_1991}, parallel \citep{hunt_fast_1995} and the fork/join \citep{kim_analysis_1989} topologies.

@@

Most models of queueing systems assume time-invariant inter-arrival and service processes. In practice, many real-world queues have inter-arrival processes which are strongly time-dependent, such as: call-centers \citep{weinberg2007bayesian, brown2005statistical}, airport runways \citep{koopman1972air} and hospitals \citep{brahimi_queueing_1991}. In the case of the $M/M/1$ queue, we can adapt the notation to $M(t)/M(t)/1$ to represent exponential  processes where parameters $\lambda(t)$ and $\mu(t)$ change with time. Such queueing systems are referred to as dynamic queueing systems.

@@

QDC can be considered as a multiserver extension to an algorithm presented by \citet{lindley_theory_1952}. For a single server queueing system, the departure time of the $i$th customer is: $d_i = \max{(a_i, d_{i-1})} + s_i$, since the customer either waits for a server or the server waits for a customer. The algorithm (not the paper) was, surprisingly, not extended to multiserver systems until \citet{krivulin_recursive_1994}. However with each new customer $i$ the algorithm must search a growing $i+1$ length vector. This algorithm, therefore, scales poorly, with computational complexity $O(n^2)$, where $n$ is number of customers. \citet{kin_generalized_2010} adapted the original algorithm of \citet{kiefer_theory_1955} to an $O(nK)$ algorithm for multiserver tandem queues with blocking, that is $G/G/K/C$ queueing systems where $C$ is the maximum capacity number of customers in the queueing systems.

QDC can also be viewed as a computationally efficient solution to the set of equations presented in \citet[pg. 259]{sutton_bayesian_2011} for FCFS queueing systems. There is a single queue served by a fixed number of $K$ servers. The $i$th customer observes a set of times $b_i = \{ b_{ik} | k \in 1:K \}$ which represents the times when each server will next be available. The customer $i$ selects the earliest available server $p_i = \text{argmin}(b_i)$ from $b_i$. The departure time for the $i$th customer is, therefore, $d_i = \max({a_i, b_{p_i}}) + s_i$, since the server must wait for the customer or the customer must wait for the server. The QDC algorithm for a fixed number of servers (Algorithm \ref{alg:fixed_servers}) pre-sorts the arrival times. Rather than assigning a $b_i$ for each customer $i$ to form the matrix $\mathbf{b} \in \mathbf{M}^{n \times K}$, QDC considers $\mathbf{b}$ as a continually updated $K$ length vector representing the state of the system.

@@

This algorithm is simple and computationally efficient. At each iteration of the loop, we need only search $\mathbf{b}$, a $K$ length vector for the minimum element in code line \ref{alg_line:search}. In the language of DES, we consider $\mathbf{b}$ as the system state and $\mathbf{a}$ as the event list, which are all determined events. This differs from conventional DES approaches to modelling queueing systems where the queue length is the system state, and both $\mathbf{a}$ and $\mathbf{d}$ constitute the event list, where the events of $\mathbf{a}$ are determined and the events of $\mathbf{d}$ are continually updated and therefore contingent.

@@

We close server $k$ by writing an $\infty$ symbol to $b_k$ ensuring that no customer can use that server. If the server needs to be open again at time $t$, we write $t$ to $b_k$ allowing customers to use that server. Since $\mathbf{x}$ now corresponds to changes in $\mathbf{b}$, it is part of the event list along with $\mathbf{a}$. The entire event list is still determined and need not be updated mid-simulation.

@@

This algorithm can simulate queues of form $G(t)/G(t)/K(t)/\infty/n/FCFS$, where $K(t)$ refers to the number of open servers changing with time. As mentioned previously this algorithm is subject to Condition \ref{eq:condition}. This condition is not overly restrictive if we consider realistic systems with few changes in $K$. The recorded server allocations $\mathbf{p} = (p_1, \cdots, p_n)$ may not reflect the real system since Algorithm \ref{alg:changing_servers} does not allow the user to specify exactly which servers are open in each epoch, only how many are open and closed. If this output is needed or in cases where Condition \ref{eq:condition} does not hold, we must use the less computationally efficient but more general unconditional algorithm below.

@@

If Condition \ref{eq:condition} does not hold or if otherwise, we wish to control exactly which servers are open at what time then we must use a less computationally efficient algorithm (Algorithm \ref{alg:changing_servers_unconditional}). Each server $k$ has its own partition of $L_k$ knot locations $\mathbf{x_k} = (x_{k,1}, \cdots, x_{k,L_k} ) \in \mathbb{R}^{L_k}_{+}$ and each $\mathbf{y_k} = (y_{k,1}, \cdots, y_{k,L_k+1} )$ is an alternating sequence of 0 and 1s of length $L_k$ indicating whether the server is open or closed respectively for the associated epoch. The vector $\mathbf{c}$ is used slightly differently to how it is used in \citet{sutton_bayesian_2011}. We use it to represent the time at which each server is next available for the current customer $i$, given the current system state $\mathbf{b}$. It is the output of the \code{next\_fun} function.

@@

This algorithm can simulate queueing systems of form $G(t)/G(t)/K(t)/\infty/n/FCFS$, where $K(t)$ refers to the number of open servers changing with time. In addition, we can specify which particular servers are available when, not just how many and we are not bound by Condition \ref{eq:condition}. Once again we note that $\mathbf{b}$ can be considered as the system state and the event list is formed by $\mathbf{a}$ and the elements of $\mathbf{\underline{x}}$. This function can be called with the \code{queue\_step} function in \pkg{queuecomputer} by supplying a \code{server.list} object to the \code{servers} argument. For the rest of this paper we focus on Algorithms \ref{alg:fixed_servers} and \ref{alg:changing_servers} for their relative conceptual simplicity and computational efficiency.

@@

With the algorithms so far presented, we can simulate from a very general set of queueing systems $G(t)/G(t)/K(t)/\infty/n/FCFS$ in a computationally efficient manner. In contrast to the algorithm of \cite{kin_generalized_2010}, the state vector $\mathbf{b}$ is written over in each iteration. The memory usage for QDC, therefore, scales with $O(n)$ rather than $O(nK)$.

@@

If the last element of $\mathbf{y}$ is zero, it is possible that some customers will never be served, this is the ``Missed customers'' output. The performance measures that follow are the mean waiting time $\bar{w}$, the mean response time $\bar{r} = d - a$, the observed utilization factor $\bar{B}/K$, the mean queue length and the mean number of customers in the system respectively. The utilization factor $\bar{B}/K$ takes into account the changing number of open servers $K(t)$ where Algorithm \ref{alg:changing_servers} is used. We now explain the implementation details of package.

@@

The \code{queue\_step} calls the more primitive \code{queue} function which is a wrapper for S3 methods which implement Algorithms \ref{alg:fixed_servers}, \ref{alg:changing_servers} or \ref{alg:changing_servers_unconditional} depending on the class of the object supplied to the \code{server} argument of \code{queue\_step}. If \code{class(server)} is numeric, then \code{queue} runs Algorithm \ref{alg:fixed_servers}, if it is a \code{server.stepfun} then \code{queue} runs Algorithm \ref{alg:changing_servers}, if it is a \code{server.list} then \code{queue} runs Algorithm \ref{alg:changing_servers_unconditional}. The \code{queue} function computes departure times $\mathbf{d}$ and server allocations $\mathbf{p}$ and the \code{queue\_step} function adds additional output such as waiting times and queue lengths which are used in summary and plot methods.

To simulate fork/join networks, the \pkg{queuecomputer} function \code{wait\_step} provides a simple wrapper to the \pkg{base} function \code{pmax.int}, this function computes the maximum of each row for a set of two equal length numeric vectors. The vectors represent the departure times for each subjob and the departure time for the entire job is the maximum of each subjob.

@@

To demonstrate the validity of the algorithm we consider a $M/M/2/\infty/1000/FCFS$ queue. If QDC is valid for any $M/M/K$ queueing system, then it is valid for any $G(t)/G(t)/K$ queueing system. This is because any non-zero $\mathbf{(a,s)}$ could conceivably come from two exponential distributions, even if the probability of the particular realization is vanishingly small. We replicate exact departure times computed with the \pkg{simmer} and \pkg{simpy} packages using \pkg{queuecomputer}. First, we generate $\mathbf{a}$ and $\mathbf{s}$ to be used as input to all three packages.

@@

We now input these objects into the three scripts using \pkg{queuecomputer}, \pkg{simmer}, or \pkg{simpy}. First, we run the \pkg{queuecomputer} script. The \code{queuecomputer_output} object is sorted in ascending order so that the departure times can be compared to the DES packages.

@@

Clearly, QDC and its implementation \pkg{queuecomputer} are a more computationally efficient way to simulate queueing systems of the form $G(t)/G(t)/K/\infty/M/FCFS$ than conventional DES algorithms implemented by \pkg{simpy} and \pkg{simmer}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width = 0.7\textwidth]{figures/benchmark_083.pdf}
  \caption{Computation time in milliseconds for varying numbers of passengers for each DES/queueing package. Each package returns exactly the same set of departure times since the same arrival and service times are supplied. The computation time reported here is the median time of 100 runs for each number of customers and each package. Intel (R) Core(TM) i7-6700
CPU @ 3.40GHz running Debian GNU/Linux.
 }
  \label{fig:bm_numberofpassengers}
\end{figure}

@@

We demonstrate \pkg{queuecomputer} by simulating a call center. The arrival time for each customer is the time that they called, and the service time is how long it takes for their problem to be resolved once they reach an available customer service representative. Let's assume that the customers arrive by a homogeneous Poisson process over the course of the day.

@@

Dashawn, however, had to wait for an available server, since he arrived after the first two customers arrived but before the first two customers departed. He must wait until one of these customers departs before he can be served. We add the departure time of the first customer (Beatriz) to his service time to compute his departure time.

@@

So the first two customers had no waiting time, but Dashawn had to wait for an available server. We can compute the waiting times for all three customers in this manner:

@@

The package integrates naturally with the popular data manipulation \proglang{R} package \pkg{dplyr} \citep{Rpkg_dplyr}. We demonstrate how to integrate \pkg{queuecomputer} and \pkg{dplyr} with a more complex Airport Terminal than before (Figure~\ref{fig:larger_airport}). Passengers from a set of 120 flights disembark at the arrivals concourse and proceed through immigration using either the ``smart gate" or the ``manual gate" route, we therefore have two queues in parallel. The route taken (smart gate or manual gate) by each passenger is predetermined, but the server used by the passenger within these separate queueing systems is not.

Their bags are unloaded from the flights and proceed to the baggage hall with a delay, the division of a passengers and bags is a fork/join network. The bags and passengers are forked at the arrival concourse and joined at the baggage hall. After immigration, the passengers proceed on to the baggage hall where they pick up their bags.

\begin{figure}[!htb]
\centering
\includegraphics[width = 0.8\textwidth]{figures/Larger_airport.pdf}
\caption{Diagram of larger airport scenario, there are 120 flights in total and two multiserver queueing systems operate in parallel. Passengers are preassigned to travel through either the ``manual" or ``smart gate" route through immigration. The passengers and bags are ``forked'' when each aircraft arrives are then ``joined'' at the baggage hall. }
\label{fig:larger_airport}
\end{figure}
